Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-04-13 ribose-paper.postfeedback.RMD
misc changes

---
title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
bibliography: resources/ribose-paper.bib
csl: resources/jap.csl
output:
  word_document:
    reference_docx: reference-template-02.docx
editor_options: 
  chunk_output_type: console
---

@@

@@

Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth. The eligibility criteria were non-smokers and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training. Lean mass (Tab 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

@@

The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Fig 1A). Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Fig 1A).
<!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo. Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo. Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway). A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs. ~0.3g Stevia rebaudiana extract). To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e. on days 1-2, 3-4, etc.). To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.

@@

## Figure/table text

 **Table 1:**  Baseline characteristics. Age = mean age in years, Height = mean height in cm, Weight and Lean mass = mean in kg, Iso 60/240 = mean unilateral isokinetic knee extension peak torque in Nm at 60 d/s and 240 d/s, Isom = mean unilateral isometric knee extension peak torque in Nm.

**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6.

@@

>>>>

---
title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
author: "Kristian Lian^1^, Sara<sup>1</sup>, Håvard Hamarsland<sup>1</sup>, Daniel Hammarström<sup>1</sup>, Stian Ellefsen<sup>1</sup>"
bibliography: resources/ribose-paper.bib
csl: resources/jap.csl
output:
  word_document:
    reference_docx: resources/reference-template-02.docx
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output"
    )
  })
---

@@

## Abstract

### Affiliations
1: Inland Norway University of Applied Sciences, Section for , Adress

### Correspondance
Kristian Lian kristian.lian@inn.no

Stian Ellefsen stian.ellefsen@inn.no

### ORCID

Kristian Lian
<a href="https://orcid.org/0000-0002-5169-5960">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
Daniel Hammarström
<a href="https://orcid.org/0000-0001-8360-2100">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>

<a href="https://orcid.org/0000-0003-0467-2018">
Stian Ellefsen
</a>,
">
Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774

</a>

\newpage

@@

Sixteen healthy male and female participants (20-33 years, Table 1) were recruited to the study through social media advertisement and word of mouth. The eligibility criteria were non-smokers and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training. Lean mass (Table 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

@@

The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Figure 1A). Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Fig 1A).
<!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo. Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo. Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway). A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs. ~0.3g Stevia rebaudiana extract). To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e. on days 1-2, 3-4, etc.). To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.

@@

## Figure captions

@@

## Tables

```{r, echo = FALSE}

###### Baseline tab
#
#
## Author: KL
#
#
## Purpose:

## Packages
library(readxl); library(tidyverse); library(flextable)

## Data

leg.dat <- read_excel("./data/dxa/ribose_dxaprleg.xlsx")

dxa.dat <- read_excel("./data/dxa/ribose_dxa.xlsx")

humac.clean2 <- readRDS("./data/data-gen/humac/humac.clean2.RDS")

str.index <- readRDS("./data/data-gen/humac/str.index.RDS")

code <- read_excel("./data/sup.xlsx")

## DXA

# Counts the "n" of females and males, added in the last step before 
# making a flextable
gendercount <- dxa.dat %>%
  count(sex) %>%
  mutate(sex = if_else(sex == "female",
                       "Female",
                       if_else(sex == "male",
                               "Male", sex)))

# Summarising and prepping dxa data
dxa.tab <- dxa.dat %>%
  select(subject, sex, age, height, weight, leanmassg) %>%
  mutate(leanmass = leanmassg/1000) %>%
  group_by(sex) %>%
  summarise(mean.a = mean(age),
            sd.a = sd(age),
            mean.h = mean(height),
            sd.h = sd(height),
            mean.w = mean(weight),
            sd.w = sd(weight),
            mean.lm = mean(leanmass),
            sd.lm = sd(leanmass)) %>%
  mutate(age = paste0(sprintf("%.1f", mean.a), " (", sprintf("%.1f",sd.a), ")"),
         height = paste0(sprintf("%.1f",mean.h), " (",sprintf("%.1f",sd.h), ")"),
         weight = paste0(sprintf("%.1f",mean.w), " (", sprintf("%.1f",sd.w), ")"),
         lm = paste0(sprintf("%.1f",mean.lm), " (", sprintf("%.1f",sd.lm), ")")) %>%
  mutate(sex = if_else(sex == "female",
                       "Female",
                       if_else(sex == "male",
                               "Male", sex))) %>%
  select(sex, age, height, weight, lm) %>%
  pivot_longer(values_to = "mean",
               names_to = "measure",
               cols = c(age:lm))

# Gets gender per subject, to mark the subjects with gender in the humac 
# data

gender <- dxa.dat %>%
  select(subject, sex)

# Summarising and prepping humac data
humac.tab <- humac.clean2 %>%
  full_join(gender) %>%
  select(subject, sex, time, acute, measure = test, peak.torque) %>%
  filter(time == "baseline" & acute == "rest") %>%
  group_by(sex, measure) %>%
  summarise(mean.pt = mean(peak.torque),
            sd.pt = sd(peak.torque)) %>%
  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
  select(sex, measure, mean) %>%
  mutate(sex = if_else(sex == "female",
                       "Female",
                       if_else(sex == "male",
                               "Male", sex)))

# Joining the data frames to be used in the flextable

dxa.tab %>%
  full_join(humac.tab) %>%
  pivot_wider(values_from = mean,
              names_from = measure) %>%
  full_join(gendercount) %>%
  select(sex, n, age, height, weight, lm, isok.60, isok.240, isom) %>%

  ## Make it to flextable
  flextable() %>%
  ## Change headers
  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header", 
          value = c(as_paragraph("Sex"),
                    as_paragraph("n"), 
                    as_paragraph("Age (yrs)"), 
                    as_paragraph("Stature (cm)"), 
                    as_paragraph("Body mass (kg)"),
                    as_paragraph("Lean mass (kg)"),
                    as_paragraph("60\U00BA sec", as_sup("-1")), 
                    as_paragraph("240\U00BA sec", as_sup("-1")), 
                    as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
  add_body_row(top = FALSE, value = as_paragraph("Values are mean and (SD)"), colwidths = 9) %>%
  set_caption("Table 1. Participant characteristics")
